Efficient Calculation of Polynomial Features on Sparse Matrices

### 2016-02-13 paper/workshop/sparse_poly.tex
removed N from big O since its for a vector
swapped a word

Polynomial feature expansion is a tool long used in statistics for approximating nonlinear functions (see \cite{gergonne1974application, smith1918standard}).
While their use is widespread, the authors of this work are unaware of any improvements made to their calculation efficiency.
In this work we provide an algorithm for calculating polynomial features for a row vector $\vec{x}$ of a $N \times D$ matrix with time and space complexity $O(d^kD^k)$ where $k$ is the polynomial order and $d$ is the density of $\vec{x}$.
The density of a vector is the percent of elements that are nonzero, so $0 \le d \le 1$.
The standard algorithm has time and space complexity $O(D^kN)$, so the added factor of $d^k$ represents a significant reduction in time.
The algorithm does not require the densification of the matrix, e.i. the matrix remains in sparse form, so the space complexity is also $O(d^kD^kN)$ as opposed to $O(D^kN)$.

@@

While the idea is straightforward, there is yet one unaddressed caveat.
Given a set of nonzero columns whose corresponding components were multiplied to produce a polynomial feature, where in the augmented polynomial vector does the result of the product belong?
To adress this, we give a bijective mapping from the set of possible column index combinations of order 2 to $k$ onto the column index space of the polynomial feature matrix.
More precisely, this the mapping is of the following form:

>>>>

Polynomial feature expansion is a tool long used in statistics for approximating nonlinear functions (see \cite{gergonne1974application, smith1918standard}).
While their use is widespread, the authors of this work are unaware of any improvements made to their calculation efficiency.
In this work we provide an algorithm for calculating polynomial features for a row vector $\vec{x}$ of a $N \times D$ matrix with time and space complexity $O(d^kD^k)$ where $k$ is the polynomial order and $d$ is the density of $\vec{x}$.
The density of a vector is the percent of elements that are nonzero, so $0 \le d \le 1$.
The standard algorithm has time and space complexity $O(D^kN)$, so the added factor of $d^k$ represents a significant reduction in time.
The algorithm does not require the densification of the matrix, e.i. the matrix remains in sparse form, so the space complexity is also $O(d^kD^k)$ as opposed to $O(D^k)$.

@@

While the idea is straightforward, there is yet one unaddressed caveat.
Given a set of columns whose corresponding nonzero components were multiplied to produce a polynomial feature, where in the augmented polynomial vector does the result of the product belong?
To adress this, we give a bijective mapping from the set of possible column index combinations of order 2 to $k$ onto the column index space of the polynomial feature matrix.
More precisely, this the mapping is of the following form:
